---
title: "Databases w/ R"
output: html_notebook
---

Borrowed and adapted from (https://github.com/edgararuiz/db_best_practices_2018)
Uses environnment from [rstudio::conf 2018](https://github.com/rstudio/bigdataclass2018)

```{r}
library(DBI)
library(odbc)

# con <- dbConnect(odbc(), 
#                       Driver = "SQL Server", 
#                       Server = "localhost\\SQLEXPRESS", 
#                       Database = "master", 
#                       Trusted_Connection = "True")

library(odbc)
con <- dbConnect(odbc::odbc(), "Postgres Dev")
```


```{r}
# dbGetQuery(con, "Select Top 10 * From datawarehouse.flight")
dbGetQuery(con, "Select * From datawarehouse.flight LIMIT 10")
```

```{sql, connection = con}
-- Select top 10 * from "flights"
Select * From datawarehouse.flight LIMIT 10
```

```{sql, connection = con, output.var = "my_var"}
Select * From datawarehouse.flight LIMIT 10
```
```{r}
my_var
```


## dplyr & dbplyr

```{r}
library(dplyr)
library(dbplyr)

```


```{r}
flights <- tbl(con, in_schema("datawarehouse", "flight"))
```

```{r}
object.size(flights)
```

```{r}
flights

```


```{r}
flights %>%
  tally()
```

```{r}
flights %>%
  group_by(month) %>%
  tally() 
```

### show_query()

```{r}
flights %>%
  group_by(dest) %>%
  tally() %>%
  show_query()
```

```{r}
complex_example <- flights %>%
  filter(distance > 100) %>%
  group_by(uniquecarrier, dest, month) %>%
  tally() %>%
  filter(n > 10) %>%
  group_by(uniquecarrier, dest) %>%
  summarize(n_avg = mean(n, na.rm = T) %>% as.integer()) %>%
  arrange(desc(n_avg))

complex_example
```
```{r}
complex_example %>% show_query()
```


### join tables

```{r}
#airports <- tbl(con, "airports")
airports <- tbl(con, in_schema("datawarehouse", "airport"))
```

```{r}
new_flights <- flights %>%
  left_join(airports, by = c("dest" = "airport"))

new_flights
```

```{r}
new_flights %>%
  group_by(state) %>%
  tally() %>%
  arrange(desc(n))
```

### collect()

```{r}
by_location <- new_flights %>%
  group_by(dest, long, lat) %>%
  tally() %>%
  collect() %>%
  mutate(n = as.numeric(n))

by_location 
```



### plotting

```{r, fig.width = 10, fig.height = 5}
library(ggplot2) 

by_location %>%
  ggplot() +
  geom_point(aes(x = long, y= lat, size = n, color = n), alpha = 0.5)

```

### Direct insights

```{r}
new_flights %>%
  group_by(dest, airportname) %>%
  tally() %>%
  arrange(desc(n)) %>%  # desc() inside arrange() get the top counts
  head(10)

```

### pull() command

```{r}
top_names <- new_flights %>%
  group_by(dest) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(10) %>% 
  collect() %>%
  pull("dest")

top_names
```

### Take advantage of dplyr's laziness

```{r}
top_flights <- new_flights %>%
  filter(dest %in% top_names)

```

Use `View()` to preview some of the data in the familiar data viewer

```{r}
top_flights %>%
  head(1000) %>%
  collect() %>%
  View("top_flights")
```

Easily transition from data manipulation to plotting, without needing to import the data

```{r}
top_flights %>%
  group_by(month) %>%
  tally() %>%
  collect() %>%
  mutate(n = as.integer(n)) %>%
  ggplot() +
    geom_area(aes(x = month , y = n), fill = "blue", alpha = 0.5)
  
```


Thanks to the `tidyverse` approach, an entirly new plot can be built by adding one single line, and the data is still summarized inside the database

```{r}
top_flights %>%
  filter(distance > 2000) %>%  # Added 
  group_by(month) %>%
  tally() %>%
  collect() %>%
  mutate(n = as.integer(n)) %>%
  ggplot() +
    geom_area(aes(x = month , y = n), fill = "blue", alpha = 0.5)
  
```

### Histograms

```{r}
library(dbplot)

flights %>%
  dbplot_histogram(arrdelay, binwidth = 50)

```
## Temporary tables

```{r}
temp_flights <- new_flights %>%
  group_by(dest) %>%
  tally() 
```

```{r}
temp_flights <- copy_to(con, temp_flights, "temp_flights")
```


```{r}
temp_flights %>%
  arrange(desc(n))
```

```{r}
dbDisconnect(con)
```




